### 2021-05-14 paper/paper.tex
paper: start of paper

\documentclass[a4paper, 10pt]{article}

\usepackage[colorlinks]{hyperref}
\usepackage{microtype}

\title{All watched over by machines of loving grace}
\author{Dominic P. Mulligan}
\date{Systems Research Group, Arm Research, Cambridge}

\begin{document}

\maketitle

\begin{abstract}
To enforce system-wide security policies, operating systems implement a self-contained \emph{kernel} which serves as the root of all trust within the computer system.
The kernel is the sole component able to directly interact with hardware, and provides services to executing user-space software, for example the abstraction of a hierarchical filesystem.
User-space code resides outside of the kernel and interacts with it through a defined \emph{ABI}.

To enforce logical soundness, many interactive theorem proving systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
The kernel is the sole system component able to authenticate a proof as legitimate.
Untrusted code resides outside of the kernel and interacts with it through a defined \emph{API}.

In this paper, we explore the consequences of structuring proof-checking software as \emph{supervisory software}.
We develop a proof-checking kernel for Gordon's HOL---an expressive typed higher-order logic---that is exposed to untrusted code through a defined ABI.
Kernel objects are protected---and important invariants are maintained---through machine-oriented notions of separation, akin to ``privilege rings'' or ``exception levels'' on typical commodity hardware, rather than programming language module boundaries.

We find that this structuring has some interesting consequences.
The typical proof-checker's \emph{meta-language}---usually some derivative of Milner's ML, privileged amongst all programming languages---is absent: any programming language able to produce code that can interact with the kernel's binary interface can be used to guide proof search, with no risk of logical unsoundness.
Moreover, the proof-checking kernel may expose functionality to untrusted code that is typical of an operating system, for example the abstraction of a hierarchical filesystem.
Filesystem access may then be conditioned on the ability to \emph{prove} some property of the dynamic system state at the point of access, allowing the kernel to enforce a range of expressive runtime security and correctness policies.
In short, untrusted code executing under the supervision of our kernel becomes \emph{proof-generating code}, exhibiting aspects of both static and dynamic (run-time) verification.

\end{abstract}

\section{Introduction}
\label{sect.introduction}

This paper is concerned with the intersection of the foundations of mathematics and systems-software design.

Modern interactive theorem proving software---such as the Isabelle, Coq, and HOL Light systems---implement a small proof-checking \emph{kernel}.
This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically small and self-contained, making it suitable for manual auditing.
By design, the kernel is the only component within the system that need be trusted by a skeptic, as ultimately all proofs constructed by the system, no matter how large or complex, are eventually passed through the kernel for authentication.
This implementation strategy---commonly known as the LCF-style, after Milner's LCF system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
That is, LCF is a design pattern that helps to write theorem proving software that remains \emph{sound}.

Shifting focus, we see an interesting series of analogies.
Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a small \emph{kernel}.
This kernel is the sole component on the compute system that has unfettered access to hardware.

\section{Kernel design}
\label{sect.kernel.design}

\section{Proof-space support libraries}
\label{sect.proof-space.support.libraries}

\end{document}